Regression Cube Analysis of Cohort Study Data

### 2015-03-25 draft.tex
added more complex revisions of Kai and Bernhard

\maketitle
\com{Bernhard: Comment on prior occurrences of "Regression Cube" and how we differ}
%%
Epidemiology aims to characterize health and disease conditions in defined populations (\emph{cohorts}).
%%
Insights about risk factors allow to characterize disease-specific high-risk groups and act as important diagnostic key figures \cite{Fletcher2012}.
%%
They can also be used to give recommendations regarding a healthy lifestyle and provide information about wide-spread diseases.
%%
In the standard workflow, physicians translate observations into hypotheses, which are depicted using epidemiological features and then assessed using regression analyses.

@@

Our contributions are:
\begin{itemize}
	\item An overview visualization technique representing feature interactions using target features.
	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
	\item Freely adjustable regression formulas provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
	\item Analysis of confounding variables by providing comparison views between different formula results
	\item The open and web-based approach of the system allows for analysis of any data using the presented method.
\end{itemize}

@@

\subsection{Epidemiological Data} \label{sec:EpidemiologicalData}
%%
\com{Bernhard: Due to overlaps to VIS'14 paper in this section, this can be shortened. E.g. "More discussion of data types and characteristics can be found in [...]"}
%%
\emph{Cohort studies} are a major epidemiological tool for gathering epidemiological features.
%%
They yield a highly heterogenous and incomplete information space.
%%
Subject data are collected either towards a specific disease or with the widest range possible.
%%
The latter allows the data set to be assessed towards different diseases.
%%
The feature space comprises information about lifestyle, somatometric variables, medical parameters, genetic data as well as medical image data.
These features are derived through different modalities, such as questionnaires or medical examinations or laboratory analyses.
%%
Many features are sparse, such as follow up questions about a medication or treatment of a certain disease.
%%
Other features are exclusive for a sub-group, such as women-specific questions, e.g., number of born children or period status.

\com{Restriction of amount of data because it must not be triangulated and because the ethics committees are very restrictive}

\paragraph{Data Types.}
%%
Medical status variables or lifestyle factors are often of \emph{dichotomous} (binary) type.
%%
The data space comprises continuous variables (somatometric variables, such as body weight or BMI or laboratory values) as well as categorical variables (e.g. graduation type).
%%
The data heterogeneity has to be taken into account when the analysis method is chosen.
%%
Continuous data are often discretized (e.g. 10 year steps for age) to equalize the variable types and to simplify the method selection.
%%
This is however avoided if possible, since it reduces the information space and also introduces a new information bias, as assumptions are modeled through the discretization.
%%

\paragraph{Image Data.}
\com{Can probably be cut (reference to Klaus' Paper for more details on this) (see Bernhards comment above)}
%%
Since the Rotterdam study, many modern cohort studies include medical image data.
%%
For ethical reasons, the imaging modalities usually not include ionizing radiation.
%%
The image quality is often inferior to clinical standards, which is a tradeoff between time and cost \cite{Preim2014}.
%%
These data are hard to analyze as they require segmentations to outline structures of interest.
%%
This process is prone to inter- and intra-observer variability when carried out manually.
%%
Automatic or semi-automatic solutions are a remedy for this problem, but are costly and need to be tailored to the structure of interest \cite{Toennies2015}.
%%
Image-derived variables are either dichotomous variables indicating a medical finding by a radiologist or continuous, describing the shape of segmented structures (e.g. volume or diameter).

@@

\paragraph{Regression Analysis Notation.} Regression formulas are usually denoted as follows:
%%
\com{Kai, + in this case is a regression operator including a feature. The formula should denote any operator (maybe as Var1 omega Var2).}
\begin{equation}
Dependent \sim Independent_1 + ... + Independent_n
%Dependent \sim \sum_{i=1}^n Independent_i
\label{eq:RegressionNotation}
\end{equation}
%%
The most used regression operators comprise:
%%
\begin{itemize}
	\item \fbox{$+,-$} inclusion/exclusion of the variable,
	\item $:$ inclusion of interactions between the variables (e.g. $x:y$),
	\item $*$ inclusion the variables as well as their interactions (e.g. $x*y$)
	\item $|$ (conditioning) inclusion of variable x, given y (e.g. $x|y$)
\end{itemize}
%%
The type of the target feature restricts the regression type.

@@

%\com{Regression cubes are an important tool for checking statistical resilience.}
%\com{Describe how regression cubes work (including notation).}
\com{Move this to Regression Cube part?}

@@

%% TODO: Check DLPD for correct page numbers
While the presented techniques were well received with epidemiological experts, the explanatory power towards back pain was limited, yielding a analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
%%
These measures were analyzed using a novel decision tree quality plot towards non-image data.
%%
We concluded that the model quality is not enough to characterize back pain.
\\\\
The difference of the presented approach towards our previous and other herein presented work is twofold:
\begin{enumerate}
	\item We focus on large scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
	\item The analysis is conducted towards a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
\end{enumerate}
\com{Main Difference to our and other related work}
\section{Regression Cube Analysis of Cohort Study Data}
\com{Add Data Preprocessing Section.}
%%
The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target variables.
%%
Overview visualizations of feature relationships as presented by Angelelli et al. \cite{Angelelli} are often focused on relationships between the visualized features.
%%
Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
%%
In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature.
%%
These target features often indicates the presence of the investigated disease.
%%
As described in Section \ref{sec:RegressionAnalysis}, regression analyses are the statistical tool of choice for analyzing these relationships.
%%
A regression model is based on expert knowledge, there is no unified rule on how to apply them to a given set of features, so they have to be applied with care.
%%
\subsection{Cube Description using Regression Formula Notation}
%%
Expert knowledge is introduced into regression analyses using the regression formulas.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target.
%%
%In order to calculate all combinations of two to three independent features, we can apply  \emph{Regression Cube} aims to 
%%

@@

\subsection{Analysis Workflow}
%% TODO: Anmerkung Kai einarbeiten Ja, gefällt mir auch nicht so gut. Ich würde zunächst Keims Mantra vorstellen (mit itemize) dann dazu erklären was er damit meint und was es bedeutet. Diese Mantrapunkte nennst du M1-M4 (oder ein anderen Buchstaben). Und dann würde ich bei der Erklärung deines Tools immer wieder darauf verweisen, welcher Punkt nun erfüllt ist (..., which fulfills M2) 
\com{ToDo. Include Kais revision, renaming Matnrapoints and refer to them}
\com{I'm not satisfied with the structure in this subsection, please proofread it with focus in the train of thought!}
\begin{figure}[htb]
 \centering
 \includegraphics[width=3.0in]{figures/workflow_sketch}
 \caption{
 Workflow of the analysis using \emph{Regression Cubes}.
 %%
 The analysis starts with a set of features given by the data set.
 %%
 Then, the user may specify a formula regarding to specify a hypothesis, or use a predefined formula to start a explorative analysis of the data set.
 %%
 The Regression Cube is then visualized, where the user has either the option to adjust the formula or to derive details on demand on a specific regression.
 %%
 Insights into the data yield either an adjustment of the current formula.
 %%
 By applying a difference view, regression cubes can be compared to each other.
 %%
 The analysis yields insights and hypothesis about feature relations.
 %%
 }
  \label{fig:Workflow}
\end{figure}
%%
%\com{Different Formulas work as different steps in the analysis (Fig.~\ref{fig:Workflow})}
%\com{Workflow for different analysis approaches.}
%\com{0. Select the Variables.}
%\com{1. Select the Formulas}
%\com{a) Analysis without hypothesis about the data set (z ~ x+y)}
%\com{b) Analysis without hypothesis with confounder (z ~ age)}
%\com{c) Analysis with hypothesis about the data set (Var ~ x+y+z)}
%\com{d) Use Difference View}
%\com{d) Clarify using confounder (Var ~ x)}
%\com{3. Find interesting pane.}
%\com{4. Get Details on Demand.}
%%
The analysis workflow follows the Visual Analytics (VA) Mantra of Keim et al. \cite{Keim}:

\textbf{Analyze First.} Choosing an initial regression formula triggers the \emph{Regression Cube} calculation on the given data set, filtering the dimensions of the dependent feature through the CFS algorithm.

\textbf{Show the Important.} The 3D-cube visualization acts as an overview over the whole data set.
%%
Regression models with large $R^2$ values can be spotted here fast, steering the users attention to the respective slice.

\textbf{Zoom, Filter and Analyze Further.} The slices of interest can then be analyzed using the 2D mosaic plot of the slice.

\textbf{Details on Demand.} Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g. in a hover-modal on a currently selected data point).
\\\\
As seen in Fig~\ref{fig:Workflow}, the workflow is highly iterative.
%%
Observations in the 2D-mosaic plot or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
%%
This can be carried out either to refine the current formula based on observations, or creating a new regression cube for a difference view.
%%
\paragraph{Cube Comparison.}
%%
Regression Cubes can be compared by creating difference views.
%%
One cube (formula) acts as reference.
%%
The absolute difference in $R^2$ values towards the second cube is calculated, yielding a difference cube showing only the differences between the two formulas.
%%
\com{TODO: divide using analysis approaches, e.g. Hypothesis-based \& Hypothesis-free with follow up on the formulas used?}
\paragraph{Hypothesis-Free Analysis}
%%
\paragraph{Hypothesis-Based Analysis}
%%
%%
\paragraph{Formula Types.}
%%
Input formulas reflect \emph{hypotheses} about the data.
%%
Using the operators, dynamic variables and dataset features, many different assumptions can be formulated.
%%
Here we show just a few examples on how formulas can serve different purposes.

$Z \sim X + Y$ is the default formula.
%%
This cube represents all possible combinations of two independent features towards all features in the data set, since we do not know which feature(s) are of interest.
%%
Each slice represents a different target feature.
%%
It is therefore suitable for a \emph{hypothesis-free} explorative analysis to give a general impression about relationships in the data set.

$Cancer \sim X + Y + Z + feature_1:feature_2$ is an example for the formulation of a hypothesis, where a specific feature is analyzed.
%%
% The interaction between $feature_1$ and $feature_2$ is also included in all regression formulas.
%%
This formula models the hypothesis, of a interaction between $feature_1$ and $feature_2$ (denoted with $:$) being relevant for the target feature, but it is not clear how other feature combination influence the result, hence the cube using $x$,$y$ and $z$ as independent features.
%%

$Cancer \sim X + Y + Z$ subtracted with the $R^2$ from $Cancer \sim Age$ excludes the confounding effect age has towards the target $Cancer$ feature.
%%

@@

	\item The \emph{file upload} section starting the analysis with providing a comma separated value (CSV) file.

@@

	It also allows to select a reference formula for creating a difference-cube.
	\item The \emph{cube visualization} consisting of the 2D-mosaic view as well as a 3D-representation of the whole cube.

@@

\paragraph{File-Upload and Classification.}

@@

\paragraph{Formula Editor.}
%%
After uploading the data, the user can specify a formula or use the default ($Z \sim X + Y$).
%%
Entering a formula is provided via text input.
%%
On entering the text are, a context modal displays all data set variables, as well as the available operators and their function.
%%
This allows to comprehend the function of the underlying formula for users without statistical background about regression analysis and its notation.
%%
Auto-completing input variables also simplifies the approach and also works as spell-check of variable names.
%%

@@

\subsection{Regression Cube Visualization.}
%%
The visualization and interaction with the \emph{Regression Cube} is the heart of the prototype.
%%
Results from the statistical processors are uploaded into the visualization by slice, allowing the assessment of the data as soon as parts of the calculations are finished while the rest is still in processing.

@@

\paragraph{3D-Prism as Data Mini-Map.}
%%
The 3D-cube representation acts as an overview over the whole data set, but its purpose is not to derive detailed information about data points.
%%
It serves an function similar to a mini-map, guiding the attention towards points of interest in the data, as well as giving context information about adjacent data values when using the 2D mosaic plot.
%%
\com{TODO Mantra}
It serves the purpose of \emph{Show the Results} after the \emph{Analyze First} step in Keim's VA-Mantra.
%%
The displayed prism shows values above the matrix diagonal.
%%
\\\\
%% TODO: Add reference for disadvantages of 3D information visualizations!
\paragraph{Tackling the disadvantages of 3D information visualization.}
3D-Information visualizations are often criticized for introducing occlusions and interaction problems, which often do not balance out the advantages of using the third dimension for visual mapping.
%%
We aim to minimize these problems as much as possible.
%%
Since the $R^2$ values are mapped on data point opacity, only large values are highlighted in the prism, guiding the focus to the respective slices.
%%
It also creates a very sparse representation, since the majority of regression models yield (depending on the data set and the chosen formula) low $R^2$ values.
%%
Also, the preceding correlation based feature selection reduces the information space significantly, leading to sparse cubes.
%%
Overlapping is still an issue, but this way greatly reduced in its affect to the readability of the visualization.
%%

%% INTERACTION
Rotating with the cube is restricted to y-axis, preserving the mental map to position individual features.
%%
The cube is always oriented according to the 2D-representation, allowing for an easy mental combination of the two representations.
%%
Allowing more degrees of freedom was confusing to our users and also does not add value to the visualization.
%%
We provide also a zoom functionality using the mouse wheel input.
%%
%\\\\
\paragraph{Cube Slice Selection.}
%%
In order to \emph{Zoom, Filter and Analyze Further}, the user has to navigate towards different slices of interest.
%%
We propose two ways to achieve this.
\begin{itemize}
	\item \emph{Applying the slicing metaphor from 3D volume data.}
	%%
	In medical volume data renderings, slicing views very common to view details on a selected plane in the scene.
	%%
	We employ this technique for selecting cube slices (e.g. by moving a plane via vertical mouse input while pressing the right mouse button).
	%%
	We, however, still display the whole 3D-object instead of cutting away information towards the slice position.
	%%
	\item \emph{Selecting the slice using a dropdown menu} provides fast access to plane selections when the user already knows the slices of interest.
	%%
\end{itemize}
%%
The currently selected slice is denoted with a semi-transparent gray plane.
%%
The space available from visualizing only the prism generated from the upper half of the cube diagonal is used to display information about the current selected plane.
%%
the values are projected on this plane to give an overlapping-free view on the data points, which makes it easier to identify the current slice.
%%
%It also allows for a fast slicing through the cube and assessing the information.
%%
%\com{display of skewed cube, combination of 2D to 3D, navigating between the two views, projection of current slice to the 3D-Cube, details on demand}
%\\\\
\paragraph{2D Mosaic Plot Slice Visualization.}
\com{Add Figure reference!}
%%
The 2D mosaic visualization shows all values below the matrix diagonal of the current slice, creating optical equivalence towards the 3D cube.
%%
To reduce visual clutter, the 2D view only shows dimensions, which are retrieved through the correlation based feature selection.
%%
The free space above the matrix diagonal is used to display the 3D cube.
%%

@@

\section{Summary and Conclusion}
\com{Rewrite, focus on precise definition on what was done and why it is awesome.}
We presented a technique for knowledge discovery in cohort study data sets with user-defined target features.
%%
\emph{Regression Cubes} are based on eponymous regression models, which allow to model domain knowledge about the data (e.g. influencing features on a disease) using formulas.
%%
Alternatively, \emph{Regression Cubes} can be calculated using default formulas to assess the explanatory power towards each feature in the data set.
%%
These two approaches allow both for a \emph{hypothesis-free} and \emph{hypothesis-based} explorative analysis.
%%
Correlation-based feature selection reduces the amount of calculations using the target feature by focusing on the important regression models.
%%
The prototype described in this work was developed using state of the art web technologies, free services and open source libraries.
%%
All code associated to this project is also open source.
%%
Two case studies revealed that the presented approach enables to reproduce knowledge extracted using decision-tree-based data mining methods on a hepatic steatosis data set, as well as producing new hypotheses by deriving insight into influencing factors on breast fat tissue in a explorative analysis.
%%

As future work, we want to introduce more regression models to the data set, which model different correlation types.
%%
We also want to extend the cube to time-dependent data by expanding the difference-cube approach.
%%

>>>>

\maketitle
%%
Epidemiology aims to characterize health and disease conditions in defined populations (\emph{cohorts}).
%%
Insights about risk factors allow to characterize disease-specific high-risk groups and act as important diagnostic key figures \cite{Fletcher2012}.
%%
They can also be used to give recommendations regarding a healthy lifestyle and provide information about wide-spread diseases.
%%
In the standard workflow, physicians translate observations into hypotheses, which are depicted using epidemiological features and then assessed using regression analyses.

@@

Our contributions are:
\begin{itemize}
	\item An overview visualization technique representing feature interactions using target features.
	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
	\item Freely adjustable regression formulas provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
	\item Analysis of confounding variables by providing comparison views between different formula results
	\item The open and web-based approach of the system allows for analysis of any data using the presented method.
\end{itemize}
%%
The term Regression Cube was used before to define for the detailed analysis of one or a small number of regression models \cite{Chan, Ahmadi}.
%%
We refer to it using as visualization technique for comparing and analyzing a large number of different regression models with different input features.
%%
\com{Bernhard, does this suffice as reasoning?}

@@

\subsection{Epidemiological Data} \label{sec:EpidemiologicalData}
%%
%\com{Bernhard: Due to overlaps to VIS'14 paper in this section, this can be shortened. E.g. "More discussion of data types and characteristics can be found in [...]"}
%%
\emph{Cohort studies} are a major epidemiological tool for gathering epidemiological features.
%%
They yield a highly heterogenous and incomplete information space.
%%
Subject data are collected either towards a specific disease or with the widest range possible.
%%
The latter allows the data set to be assessed towards different diseases.
%%
The feature space comprises information about lifestyle, somatometric variables, medical parameters, genetic data as well as medical image data.
These features are derived through different modalities, such as questionnaires or medical examinations or laboratory analyses.
%%
Many features are sparse, such as follow up questions about a medication or treatment of a certain disease.
%%
Other features are exclusive for a sub-group, such as women-specific questions, e.g., number of born children or period status.
%%
%\com{Restriction of amount of data because it must not be triangulated and because the ethics committees are very restrictive}
\\\\
%\paragraph{Data Types.}
%%
Medical status variables or lifestyle factors are often of \emph{dichotomous} (binary) type.
%%
The data space comprises continuous variables (somatometric variables, such as body weight or BMI or laboratory values) as well as categorical variables (e.g. graduation type).
%%
The data heterogeneity has to be taken into account when the analysis method is chosen.
%%
Continuous data are often discretized (e.g. 10 year steps for age) to equalize the variable types and to simplify the method selection.
%%
This is however avoided if possible, since it reduces the information space and also introduces a new information bias, as assumptions are modeled through the discretization.
%%

Medical image data is also acquired and analyzed in modern cohort studies.
%%
We incorporate image-derived features, but do not put focus on analyzing image data.
%%
More discussion of data types and characteristics can be found in \cite{Klemm2014VIS, Toennies2015, Preim2014}.

%\paragraph{Image Data.}
% \com{Can probably be cut (reference to Klaus' Paper for more details on this) (see Bernhards comment above)}
% %%
% Since the Rotterdam study, many modern cohort studies include medical image data.
% %%
% For ethical reasons, the imaging modalities usually not include ionizing radiation.
% %%
% The image quality is often inferior to clinical standards, which is a tradeoff between time and cost \cite{Preim2014}.
% %%
% These data are hard to analyze as they require segmentations to outline structures of interest.
% %%
% This process is prone to inter- and intra-observer variability when carried out manually.
% %%
% Automatic or semi-automatic solutions are a remedy for this problem, but are costly and need to be tailored to the structure of interest \cite{Toennies2015}.
% %%
% Image-derived variables are either dichotomous variables indicating a medical finding by a radiologist or continuous, describing the shape of segmented structures (e.g. volume or diameter).

@@

\paragraph{Regression Analysis Notation.} Regression formulas are usually denoted as follows:
%%
\begin{equation}
Dependent \sim Independent_1 + ... + Independent_n
%Dependent \sim \sum_{i=1}^n Independent_i
\label{eq:RegressionNotation}
\end{equation}
%%
The most used regression operators comprise:
%%
\begin{itemize}
	\item \fbox{$+,-$} inclusion/exclusion of the variable,
	\item $:$ inclusion of interactions between the variables (e.g. $x:y$),
	\item $*$ inclusion the variables as well as their interactions (e.g. $x*y$)
	\item $|$ (conditioning) inclusion of variable x, given y (e.g. $x|y$)
\end{itemize}
%%
The type of the target feature restricts the regression type.

@@

%\com{Regression cubes are an important tool for checking statistical resilience.}
%\com{Describe how regression cubes work (including notation).}
%\com{Move this to Regression Cube part?}

@@

%% TODO: Check DLPD for correct page numbers
While the presented techniques were well received with epidemiological experts, the explanatory power towards back pain was limited, yielding a analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
%%
These measures were analyzed using a novel decision tree quality plot towards non-image data.
%%
We concluded that the model quality is not enough to characterize back pain.
\\\\
The difference of the presented approach towards our previous and other herein presented work is twofold:
\begin{enumerate}
	\item We focus on large scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
	\item The analysis is conducted towards a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
\end{enumerate}
%%
\section{Regression Cube Analysis of Cohort Study Data}
% TODO: Do we need a preprocessing section?
%\com{Add Data Preprocessing Section.}
%%
The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target variables.
%%
Overview visualizations of feature relationships as presented by Angelelli et al. \cite{Angelelli} are often focused on relationships between the visualized features.
%%
Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
%%
In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature.
%%
These target features often indicates the presence of the investigated disease.
%%
As described in Section \ref{sec:RegressionAnalysis}, regression analyses are the statistical tool of choice for analyzing these relationships.
%%
A regression model is based on expert knowledge, there is no unified rule on how to apply them to a given set of features, so they have to be applied with care.
%%
\subsection{Cube Description using Regression Formula Notation}
%%
Expert knowledge is introduced into regression analyses using the regression formulas.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target.
%%
%In order to calculate all combinations of two to three independent features, we can apply  \emph{Regression Cube} aims to 
%%

@@

\subsection{Analysis Workflow}
%%
\begin{figure}[htb]
 \centering
 \includegraphics[width=3.0in]{figures/workflow_sketch}
 \caption{
 Workflow of the analysis using \emph{Regression Cubes}.
 %%
 [\textbf{1.AF}] The analysis starts with a set of features given by the data set.
 %%
 Then, the user may specify a formula regarding to specify a hypothesis, or use a predefined formula to start a explorative analysis of the data set.
 %%
 [\textbf{2.SI}] The Regression Cube is then visualized, where the user has either the option to adjust the formula or to derive details on demand on a specific regression.
 %%
 [\textbf{3.ZF}] Insights into the data yield either an adjustment of the current formula.
 %%
 By applying a difference view, regression cubes can be compared to each other.
 %%
 [\textbf{4.DD}] Details about features using the 2D-heatmap representation yields insights and hypothesis about feature relations.
 }
  \label{fig:Workflow}
\end{figure}
%%
%\com{Different Formulas work as different steps in the analysis (Fig.~\ref{fig:Workflow})}
%\com{Workflow for different analysis approaches.}
%\com{0. Select the Variables.}
%\com{1. Select the Formulas}
%\com{a) Analysis without hypothesis about the data set (z ~ x+y)}
%\com{b) Analysis without hypothesis with confounder (z ~ age)}
%\com{c) Analysis with hypothesis about the data set (Var ~ x+y+z)}
%\com{d) Use Difference View}
%\com{d) Clarify using confounder (Var ~ x)}
%\com{3. Find interesting pane.}
%\com{4. Get Details on Demand.}
%%
The analysis workflow follows the Visual Analytics (VA) Mantra of Keim et al. \cite{Keim}:

\textbf{Analyze First} [\textbf{1.AF}]. Choosing an initial regression formula triggers the \emph{Regression Cube} calculation on the given data set, filtering the dimensions of the dependent feature through the CFS algorithm.

\textbf{Show the Important} [\textbf{2.SI}]. The 3D-cube visualization acts as an overview over the whole data set.
%%
Regression models with large $R^2$ values can be spotted here fast, steering the users attention to the respective slice.

\textbf{Zoom, Filter and Analyze Further} [\textbf{3.ZF}]. The slices of interest can then be analyzed using the 2D mosaic plot of the slice.

\textbf{Details on Demand} [\textbf{4.DD}]. Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g. in a hover-modal on a currently selected data point).
\\\\
We use the squared bracket abbreviation for each step to denote the affiliation the system design section later on.
%%
As seen in Fig~\ref{fig:Workflow}, the workflow is highly iterative.
%%
Observations in the 2D-mosaic plot or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
%%
This can be carried out either to refine the current formula based on observations, or creating a new regression cube for a difference view.
%%
\paragraph{Hypothesis-Free and Hypothesis-Based Analysis.}
%%
Input formulas reflect \emph{hypotheses} about the data.
%%
Using the operators, dynamic variables and dataset features, many different assumptions can be formulated.
%%
To support \emph{hypothesis-free} analysis by providing a default formula:

$Z \sim X + Y$.
%%
This cube represents all possible combinations of two independent features towards all features in the data set, since we do not know which feature(s) are of interest.
%%
Each slice represents a different target feature.
%%
It is therefore suitable for a explorative analysis to give a general impression about relationships in the data set.

Hypotheses about the data are reflected in either how dynamic variables are related using the regression operators.
%%
Also, static features can be added for each regression formula.
%%
Here are a few examples:

%%
$Cancer \sim X + Y + Z$ is an example for the formulation of a hypothesis, where a specific feature is analyzed.
%%
All combinations of three independent features towards the target are analyzed through the cube.

$Cancer \sim X + Y + Z + feature_1:feature_2$ encodes more assumptions.
%%
This formula models the hypothesis of an interaction between $feature_1$ and $feature_2$ (denoted with $:$) being relevant for the target feature, but it is not clear how other feature combination influence the result.
%%
Therefore, the cube incorporates this interaction for all $X$,$Y$ and $Z$ values as independent features.
%%

$Cancer \sim X + Y + Z$ subtracted with the $R^2$ from $Cancer \sim Age$ excludes the confounding effect age has towards the target $Cancer$ feature.
%%
This is achieved through cube comparison.
%%

\paragraph{Cube Comparison.}
%%
Regression Cubes can be compared by creating difference views.
%%
One cube (formula) acts as reference.
%%
The absolute difference in $R^2$ values towards the second cube is calculated, yielding a difference cube showing only the differences between the two formulas.
%%
It can for example be utilized for comparing the influence of one feature towards the complete result (e.g., $Z \sim X + Y$ and $Z \sim X + Y + Income$)
%%
% \paragraph{Formula Types.}
% %%
% Input formulas reflect \emph{hypotheses} about the data.
% %%
% Using the operators, dynamic variables and dataset features, many different assumptions can be formulated.
% %%
% Here we show just a few examples on how formulas can serve different purposes.
%
% $Z \sim X + Y$ is the default formula.
% %%
% This cube represents all possible combinations of two independent features towards all features in the data set, since we do not know which feature(s) are of interest.
% %%
% Each slice represents a different target feature.
% %%
% It is therefore suitable for a \emph{hypothesis-free} explorative analysis to give a general impression about relationships in the data set.

@@

	\item The \emph{file upload} section starting the analysis with providing a comma separated value (CSV) file [1.AF].
	\item The \emph{cube visualization} consisting of the 2D-mosaic view as well as a 3D-representation of the whole cube [2.SI].

@@

	It also allows to select a reference formula for creating a difference-cube [3.ZF].

@@

\paragraph{File-Upload and Classification [1.AF].}

@@

\paragraph{Formula Editor [1.AF, 3.ZF].}
%%
After uploading the data, the user can specify a formula or use the default ($Z \sim X + Y$).
%%
Entering a formula is provided via text input.
%%
On entering the text are, a context modal displays all data set variables, as well as the available operators and their function.
%%
This allows to comprehend the function of the underlying formula for users without statistical background about regression analysis and its notation.
%%
Auto-completing input variables also simplifies the approach and also works as spell-check of variable names.
%%

@@

\subsection{Regression Cube Visualization [2.SI].}
%%
The visualization and interaction with the \emph{Regression Cube} is the heart of the prototype.
%%
Results from the statistical processors are uploaded into the visualization by slice, allowing the assessment of the data as soon as parts of the calculations are finished while the rest is still in processing.

@@

\paragraph{3D-Prism as Data Mini-Map.}
%%
The 3D-cube representation acts as an overview over the whole data set, but its purpose is not to derive detailed information about data points.
%%
It serves an function similar to a mini-map, guiding the attention towards points of interest in the data, as well as giving context information about adjacent data values when using the 2D mosaic plot.
%%
It serves the purpose of \emph{Show the Results} after the \emph{Analyze First} step in Keim's VA-Mantra.
%%
The displayed prism shows values above the matrix diagonal.
%%
%% TODO: Add reference for disadvantages of 3D information visualizations!
\paragraph{Tackling the disadvantages of 3D information visualization.}
3D-Information visualizations are often criticized for introducing occlusions and interaction problems, which often do not balance out the advantages of using the third dimension for visual mapping.
%%
We aim to minimize these problems as much as possible.
%%
Since the $R^2$ values are mapped on data point opacity, only large values are highlighted in the prism, guiding the focus to the respective slices.
%%
It also creates a very sparse representation, since the majority of regression models yield (depending on the data set and the chosen formula) low $R^2$ values.
%%
Also, the preceding correlation based feature selection reduces the information space significantly, leading to sparse cubes.
%%
Overlapping is still an issue, but this way greatly reduced in its affect to the readability of the visualization.
%%

%% INTERACTION
Rotating with the cube is restricted to y-axis, preserving the mental map to position individual features.
%%
The cube is always oriented according to the 2D-representation, allowing for an easy mental combination of the two representations.
%%
Allowing more degrees of freedom was confusing to our users and also does not add value to the visualization.
%%
We provide also a zoom functionality using the mouse wheel input.
%%
%\\\\
\paragraph{Cube Slice Selection [3.ZF].}
%%
In order to \emph{Zoom, Filter and Analyze Further}, the user has to navigate towards different slices of interest.
%%
We propose two ways to achieve this.
\begin{itemize}
	\item \emph{Applying the slicing metaphor from 3D volume data.}
	%%
	In medical volume data renderings, slicing views very common to view details on a selected plane in the scene.
	%%
	We employ this technique for selecting cube slices (e.g. by moving a plane via vertical mouse input while pressing the right mouse button).
	%%
	We, however, still display the whole 3D-object instead of cutting away information towards the slice position.
	%%
	\item \emph{Selecting the slice using a dropdown menu} provides fast access to plane selections when the user already knows the slices of interest.
	%%
\end{itemize}
%%
The currently selected slice is denoted with a semi-transparent gray plane.
%%
The space available from visualizing only the prism generated from the upper half of the cube diagonal is used to display information about the current selected plane.
%%
the values are projected on this plane to give an overlapping-free view on the data points, which makes it easier to identify the current slice.
%%
%It also allows for a fast slicing through the cube and assessing the information.
%%
%\com{display of skewed cube, combination of 2D to 3D, navigating between the two views, projection of current slice to the 3D-Cube, details on demand}
%\\\\
\paragraph{2D Mosaic Plot Slice Visualization [4.DD].}
\com{Add Figure reference!}
%%
The 2D mosaic visualization shows all values below the matrix diagonal of the current slice, creating optical equivalence towards the 3D cube.
%%
To reduce visual clutter, the 2D view only shows dimensions, which are retrieved through the correlation based feature selection.
%%
The free space above the matrix diagonal is used to display the 3D cube.
%%

@@

\section{Summary and Conclusion}
We presented a technique for knowledge discovery in cohort study data sets with user-defined target features.
%%
Dimension reduction using the target restricts the analysis to most important features.
%%
\emph{Hypothesis-free} analysis employs default regression models.
%%
Model expert knowledge using regression formulas allow for \emph{hypothesis-based} investigation.
%%
A 3D \emph{Regression Cube} allows to assess hotspots in the analysis by abstracting regression models using a quality-of-fit measure.
%%
These can then be analyzed further using the 2D mosaic plot for each cube slice.
%%
Details on demand for each model allow for detailed assessment of regression models.
%%
We successfully applied the approach to find correlations in a hepatic steatosis as well as a breast cancer data set.
%%
The method was well received by our clinical partners, triggering detailed investigations of the findings.

As next step, we want to introduce more regression types, which model different kinds of correlations.
%%
We also want to extend the cube to time-dependent data by expanding the difference-cube approach.
%%

We published all associated code and provide an analysis platform open to heterogenous data types.
%%
We believe in the power of opening up knowledge discovery to allow a heterogenous group of domain experts to derive insight into their data and support the notion of open science.
